Risks of Running Vulnerable Container Images for Data Analysis on HPC Clusters

### 2019-08-22 paper.tex
updated intro

%%%% Packages %%%%
\usepackage{siunitx}
\usepackage{algpseudocode} % Algorithmic environment
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{makecell}
\usepackage[flushleft]{threeparttable}

@@

Virtualization, which was introduced in 1960s, refers to the creation of a virtual
version of something. Its meaning has evolved and took different forms since then, for
instance hardware virtualization, containerization, desktop virtualization, etc.
Here, in this paper our main focus is on containerarization and its security concerns,
specifically in neuroimaging field.
Containers, which provide operating system (OS) level virtualization, are very
popular. This is due to the need for fast development cycles, continuous delivery,
and cost reduction of infrastructure. Moreover, demand for an isolated
environment adds on to the popularity of containers. Indeed, containers accelerate
development cycles as compared to virtual machines and provide near-native performance.
They are very lightweight, flexible, and more resource-efficient than virtual
machines. By providing OS level virtualization they
removes the overhead of having an extra OS layer.

Although Linux Containers (LXC) and other alternatives such as Rocket were already
available, container-based virtualization gained a lot of interest when Docker was
introduced in 2013~\cite{gantikow2016providing}. Docker makes container's use more straightforward. However, there
are few limitations of Docker. For example, it is not supported by shared High Performance
Computing (HPC) environments because its default configuration runs containers
as root. Secondly, Docker containers cannot be used for reproducibility as Docker
images can be updated, and hence losing specific package versions required for
reproducing scientific results. Consequently, Singularity was introduced in 2016 with the main
goal of removing drawbacks of Docker. Singularity was developed mainly to
support HPC environments and to provide reproducibility. \TG{The previous two paragraphs are too general. You should
summarize them in 2-3 sharper sentences.}

No doubt containers have gained a lot of popularity after Docker, however containers tightly
integrate with the host due to the sharing of the kernel, 
mounted file systems, and devices, which raises security
concerns. Figure \ref{fig:container-overview} provides overview of the containers architecture.
They also share three main security threats with other virtualization
techniques~\cite{gantikow2016providing}.

@@

To answer that question, we scan all images that are used on Compute
Canada clusters by neuroscience field. Scanning refers to the practice of
collecting all information about the container image, investigating it
for vulnerabilities, and finally producing a
report to summarize scanning results. Scanning is broadly categorized into
two types: static scanning and dynamic scanning. The former scanning technique
involves extracting information of an image without running it into a container,
whereas the latter approach involves launching the container from the target image, and
then scanning the running container.
Here, we focus mainly on vulnerability scanning and not on malware.
Malware is a computer software that is specially designed to attack the target computer,
whereas vulnerability is a weakness in the developer's software that can be used
by the attacker to perform unathorized actions on the target computer.
In this paper, we report all vulnerabilities that are present in
container images, specifically used by neuroimaging field, and access their
related risks for data analysis. Particularly, we aim to answer three primary
research questions:

@@

In this section we summarize a particular use case of containers, Canadian
Open Neuroscience Platform (CONP) \TG{No. In this section you describe the data and tools 
that were used in your experiments. The fact that they are used in CONP 
is a side note}. It is a national platform for sharing neuroscience
research data and aims at creating an interactive interface for neuroscientists and
clinical neuroscience. The main goal of the CONP platform is to store, process, and
distribute massive amounts of data produced by modern neuroscience.
Here, we summarize all the resources that are
used by CONP to function.

\subsection{Compute Canada Clusters}

Compute Canada is a national High Performance Computing (HPC) research platform.
It provides infrastructure and services for Advanced Research Computing (ARC).
Its goal is to bring excellence, ease, and innovation in research of Canadian
research institutions by providing storage, computing power, and required
softwares to carry out research.
It integrates with four regional partners namely ACENET, Calcul Québec, Compute
Ontario, and WestGrid for providing local support.
Compute Canada uses Slurm Workload Manager to schedule jobs and Lustre shared
file system to provide high bandwidth to its users.
Control groups are used by it to limit number of resources that can be used
by a particular user.
Compute Canada do not support Docker as its default configuration
runs containers as root. However, Compute Canada supports Singularity
containers because it removes the drawback of Docker.
\TG{Not sure if Compute Canada is used in the experiments. If not, remove and 
pitch it in the introduction that containers are nowadays used in many clusters, including 
Compute Canada. You could also refer to the full list of clusters that support singularity.}

@@

Singularity, which is introduced in 2016, provides operating system (OS) level
virtualization.
It offers mobility of compute by facilitating portable environments 
through a single image file~\cite{kurtzer2016singularity}. Once Singularity image
is created, it is hashed by using SHA256 hashing. Hence, it cannot be changed
once it is created, consequently, it can be used for reproducing the results of
scientific experiments. Singularity was introduced with the main goal of
providing support for multi-user environment. Main features provided by
Singularity are mobility of compute and reproducibility~\cite{kurtzer2017singularity}.
Mobility of compute is defined as creating and maintaining a workflow on
a local machine that can be easily ported on other hosts, Linux distributions,
and/or cloud service providers. To achieve this, Singularity packs everything
that is required for the application to run in a distributable image. Then that
image can be copied, archived, and shared. Moreover, Singularity containers
are easily portable across different versions of the C library and different kernel
implementations.
Same features of Singularity, which are used for mobility of compute,
facilitate reproducibility as well~\cite{sochat2017enhancing}.

\TG{This paragraph is too general for a method section. You might summarize this in
2-3 sentences for the introduction. Here you should talk about the engine,
i.e., how images are run, focusing on security features. Use of namespaces,
cgroups, access to devices, etc}

\TG{You should add a subsection presenting Docker.}

@@

Docker images are organized as a series of layers stacked on top of each
other and identified by unique layer identifiers. If a layer is reused
in different images it will
have the same id across images. Each layer consists of a filesystem diff
with respect to the layer below. Layers are compressed \TG{Are you
referring to file compression similar to gzip?} into a single image and can
be pushed to DockerHub. There are two ways to create a Docker image. The
first is to use an existing Docker image as a base image and then build a
new image from that one, by making changes in the filesystem, which is
saved as a new layer of the Docker image. The second way is to build a new
Docker image from scratch. Docker supports several different storage drivers.
\textit{Overlay2} is the preferred storage driver used by Docker for all currently
supported Linux versions, whereas \textit{aufs} is preferred for Docker 18.06 and
older, when using Ubuntu 14.04 on kernel 3.13 as it do not support \textit{overlay2}.
Other storage drivers that are supported by Docker include \textit{devicemapper, btrfs,
zfs, and vfs}

@@

Singularity takes snapshot, locks, and archives the developed application
\TG{what does it mean?}, which is known as Singularity image. Due to image
hashing, Singularity does not contain image layers, unlike Docker images
\TG{is it really due to hashing? why couldn't a hashed image have layers?}.
So Singularity image is a single immutable file that can be moved around as
any other file \TG{Talking about the file system used would be relevant}.
While publishing experimental results, authors can also publish Singularity
image along with its hash, which allow other researchers to verify results.

@@

\begin{figure}
  \centering
  \includegraphics[width=.7\columnwidth]{Figures/bids_table.png}
  \caption{All Container images used by BIDS apps.}
  \label{fig:bids_apps}
\end{figure}

@@

If there exist vulnerabilities in the container image, then there are two ways
to proceed, either not to run that image at all, or run only after updating the
image. Updating the image involves updating all software packages that are
inside that image. However, this option effects the reproducibility
of the image, which means that the image cannot be used to verify research
findings of other scientists. An alternative approach can be to trim unnecessary
packages from images, which in turn reduces the number of vulnerabilities
present in images. \TG{Here you should describe the method you used to update the images, linking to the 
software (script) you used to do that. 
This paragraph is too general and doesn't belong to the methods. }

@@

Anchore is an end-to-end, open-source container security platform. Anchore
is a static \TG{You should explain static vs dynamic scanning in the intro}
scanning tool that analyses container images and lists operating system
packages, unofficial packages, configuration files \TG{How about files
which are not configuration files?}, and language modules [58] \TG{Import
reference}. \TG{What is a language module?} \TG{Does it list unknown
packages, files or modules?} Anchore is a set of tools, which users can
employ for scanning container images and applying custom policies to them.
It provides visibility and transparency of the container environment.
Anchore offers services in two ways: Anchore Engine and Anchore Engine
Command Line Interface (CLI). Anchore Engine is available as a Docker
image, which is available on Docker Hub. Anchore Engine CLI provides
command line interface in addition to Anchore Engine REST API. Anchore
Engine CLI can be installed directly on the host and used to inspect
images.

@@

Vuls is an open-source vulnerability scanner for Linux and FreeBSD.
It is an agentless vulnerability scanner and uses various
vulnerability databases \TG{You should list them and explain how they differ to 
the ones used by Anchore}. Using dynamic scanning technique, it offers
remote and local scan. Remote scan involves setting up only one machine, to
which other target servers are connected via SSH. Local scan involves running
the scan on the target server itself. Further, Vuls offer fast scan, fast root scan,
and deep scan.
Fast scan is performed without root privileges, whereas fast root scan is
performed with root privileges. The latter also uses checkrestart utility to 
detect processes that are updated but not yet restarted because if they are not
restarted, updates are not effective. So fast root scan also detects these pro-
cesses. Both fast scan and fast root scan are performed offline without Internet
access, and put almost no load on the scan target server. Deep scan uses
root privileges for scanning, and is used to get a more detailed scan. It checks
changelogs, which puts more load on the target server.

>>>>

%%%% Packages %%%%
\usepackage{siunitx}
\usepackage{algpseudocode} % Algorithmic environment
\usepackage{xspace}
\usepackage{minted} % Used for JSON highlighting
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{makecell}
\usepackage[flushleft]{threeparttable}

@@

Containers, which provide operating system (OS) level virtualization, are very
popular. This is because they are very lightweight, flexible, resource-efficient,
provide fast development cycles, continuous delivery,
and cost reduction of infrastructure. OS-level virtualization gained a lot of interest when Docker was
introduced in 2013~\cite{gantikow2016providing}. However, it has security concerns due to which it
is not supported by High Performance
Computing (HPC) environments. As a result, Singularity came in 2016 with the goal of providing support
for HPC environments. It is used in many clusters including Compute Canada.
Table \ref{fig:cluster_list} illustrates the reported usage of Singularity on HPC clusters~\cite{kurtzer2017singularity}.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{Figures/cluster_list.png}
	\caption{List of HPC clusters using Singularity~\cite{kurtzer2017singularity}.}
	\label{fig:cluster_list}
\end{figure}

No doubt containers have gained a lot of popularity, however containers tightly
integrate with the host due to the sharing of the kernel, 
mounted file systems, and devices, which raises security
concerns. Figure \ref{fig:container-overview} provides overview of the containers architecture.
They also share three main security threats with other virtualization
techniques~\cite{gantikow2016providing}.

@@

To answer that question, we scan all images that are used on Compute
Canada clusters by neuroscience field. Scanning refers to the practice of
collecting all information about the container image, investigating it
for vulnerabilities, and finally producing a
report to summarize scanning results. It is broadly categorized into
two types: static scanning and dynamic scanning. The former scanning technique
involves extracting information of an image without running it into a container,
whereas the latter approach involves launching the container from the target image, and
then scanning the running container.
Here, we focus mainly on vulnerability scanning and not on malware.
Malware is a computer software that is specially designed to attack the target computer,
whereas vulnerability is a weakness in the developer's software that can be used
by the attacker to perform unathorized actions on the target computer.
In this paper, we report all vulnerabilities that are present in
container images, specifically used by neuroimaging field, and access their
related risks for data analysis. Particularly, we aim to answer three primary
research questions:

@@

As we are taking a particular use case of containers from
neuroscience field i.e Canadian Open Neuroscoence Platform (CONP) so  
in this section we summarize the data and tools 
that are used in CONP. CONP is a national platform for sharing neuroscience
research data and aims at creating an interactive interface for neuroscientists and
clinical neuroscience. The main goal of the CONP platform is to store, process, and
distribute massive amounts of data produced by modern neuroscience.
Here, we summarize all the resources that are
used by CONP to function.

@@

Singularity offers mobility of compute by facilitating portable environments 
through a single image file~\cite{kurtzer2016singularity}. Once Singularity image
is created, it is hashed by using SHA256 hashing. Hence, it cannot be changed
once it is created, consequently, it can be used for reproducing the results of
scientific experiments. Main features provided by
Singularity are mobility of compute and reproducibility~\cite{kurtzer2017singularity}.

Users can easily access files, devices, sockets, ports of the host system from
the Singularity container. By default, Singularity bind mounts /home/\$USER, /tmp, and \$PWD
into the running container. Additional directories can be specified with \textit{--bind}
to bind mount into the container. By defaut, Singularity uses as less as possible namespaces.
However namespaces can be requested if required. Users cannot escalate privileges inside
Singularty containers. This is because Singularity do not run containers as root.
Instead, a user inside Singularity container is the same user outside the container.

\subsubsection{Docker Engine}

Docker runs a software, \textit{Docker daemon}, on the host machine.
Docker daemon runs as root, and is remotely controlled through a
UNIX socket. It is responsible for launching containers, their
isolation, and spawning shells into the containers. It is also
responsible for managing Docker images i.e \textit{pull} and \textit{push}
from image repository, building images from Dockerfiles, signing them, etc..
When a Docker container runs, the container process is spawned as a
child of this root owned daemon. As Docker users directly interact with the
Docker daemon, it is possible to force Docker daemon to escalate privileges.
If a user gets \textit{root} privileges, it introduces unthinkable security risks
for multi-tenant shared enviornments. Due to this reason, Docker is not
supported by HPC environments. However, Docker containers can be imported into
Singularity and then used on HPC clusters.

@@

Docker images are organized as a series of layers stacked on top of each
other and identified by unique layer identifiers. If a layer is reused
in different images it will
have the same id across images. Each layer consists of a filesystem diff
with respect to the layer below. DockerHub uses gzip compression for
transferring layers. There are two ways to create a Docker image. The
first is to use an existing Docker image as a base image and then build a
new image from that one, by making changes in the filesystem, which is
saved as a new layer of the Docker image. The second way is to build a new
Docker image from scratch. Docker supports several different storage drivers.
\textit{Overlay2} is the preferred storage driver used by Docker for all currently
supported Linux versions, whereas \textit{aufs} is preferred for Docker 18.06 and
older, when using Ubuntu 14.04 on kernel 3.13 as it do not support \textit{overlay2}.
Other storage drivers that are supported by Docker include \textit{devicemapper, btrfs,
zfs, and vfs}

@@

Singularity takes snapshot, locks, and archives the developed application,
which is known as Singularity image.
It means that Singularity image is a single immutable file that can be moved around as
any other file. This file can be checksummed, signed, and easily
used for verification purposes. Singularity containers use Squashfs file format
by default. However, it also supports two other formats i.e \textit{sandbox} format
(which is just a chroot
directory) and \textit{writable} format (the ext3 file system).
While publishing experimental results, authors can also publish Singularity
image along with its hash, which allow other researchers to verify results.

@@

\begin{figure}
  \centering
	\includegraphics[width=\columnwidth]{Figures/bids_table.png}
  \caption{All Container images used by BIDS apps.}
  \label{fig:bids_apps}
\end{figure}

@@

If there exist vulnerabilities in the container image, then there are two ways
to proceed, either not to run that image at all, or run only after updating the
image. Updating the image involves updating all software packages that are
inside that image. However, this option effects the reproducibility
of the image, which means that the image cannot be used to verify research
findings of other scientists. An alternative approach can be to trim unnecessary
packages from images, which in turn reduces the number of vulnerabilities
present in images. To update container images we used a bash script
which is given in Figure~\ref{}.

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=0pt,
               tabsize=4]{bash}
#!/usr/bin/env bash

function die {
	echo $1
	exit 1
}

function find-pm {
		(type $1 &> /dev/null && echo "> Found $1") || \
		(echo "> $1 not found" && return 1)
}

\rm -f .updated
find-pm yum && (yum update -y && touch .updated) 
find-pm apt && (apt update -y && apt upgrade -y \
&& apt autoremove -y && apt autoclean -y \
&& touch .updated)
test -f .updated || die "Cannot find package manager \
or update failed"	       
\end{minted}
\caption{update.sh}
\label{listing:update.sh}
\end{listing}

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=0pt,
	       tabsize=4]{bash}
#!/usr/bin/env bash

set -ue

function usage {
	echo "$0 <image_list>"
	exit 1
}

if [ $# -lt 1 ]
then
	usage
fi

for image in $*
do
	echo -n "Updating image ${image}... "
	docker run -e /bin/bash -v $PWD:$PWD -w \
	$PWD ${image} ./update.sh &>${image}.log \
	&& echo "[ OK ]"
	echo -n "Commiting updated image... "
	ID=$(docker ps -a | head -2 | awk \
	'$1!="CONTAINER" {print $1}')
	docker commit ${ID} updated_${image} \
	&& echo "[ OK ]"
done
\end{minted}
\caption{update-images.sh}
\label{listing:update-images.sh}
\end{listing}

@@

Anchore is an end-to-end, open-source container security platform. Anchore
is a static
scanning tool that analyses container images and lists OS
packages, non-OS packages (Python, Java, Gem, and Npm), and files.
Anchore is a set of tools, which users can
employ for scanning container images and applying custom policies to them.
It provides visibility and transparency of the container environment.
Anchore offers services in two ways: Anchore Engine and Anchore Engine
Command Line Interface (CLI). Anchore Engine is available as a Docker
image, which is available on Docker Hub. Anchore Engine CLI provides
command line interface in addition to Anchore Engine REST API. Anchore
Engine CLI can be installed directly on the host and used to inspect
images.

@@

Vuls is an open-source vulnerability scanner for Linux and FreeBSD.
It is an agentless vulnerability scanner and uses various
vulnerability databases. Using dynamic scanning technique, it offers
remote and local scan. Remote scan involves setting up only one machine, to
which other target servers are connected via SSH. Local scan involves running
the scan on the target server itself. Further, Vuls offer fast scan, fast root scan,
and deep scan.
Fast scan is performed without root privileges, whereas fast root scan is
performed with root privileges. The latter also uses checkrestart utility to 
detect processes that are updated but not yet restarted because if they are not
restarted, updates are not effective. So fast root scan also detects these pro-
cesses. Both fast scan and fast root scan are performed offline without Internet
access, and put almost no load on the scan target server. Deep scan uses
root privileges for scanning, and is used to get a more detailed scan. It checks
changelogs, which puts more load on the target server.
